TGI Spinal Analysis Code Experiment 1

### 2023-07-12 Manuscript.Rmd
updates

## Thermosensory and burning components of TGI perception are spinally mediated 
```{r, warning = F, message = F, echo = F,include=FALSE}
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(boot, broom, broom.mixed, brms, car, caret, corrplot, DHARMa, gamlss,
                 ggeffects, gghalves, ggpol, ggpubr, glmmTMB, groupdata2, lme4, lmerTest, 
                 modelsummary, MuMIn, polycor, RColorBrewer, rcompanion, reshape2, Rmisc,
                 rsample, rstatix, simr, tibble, tidyverse, wesanderson, patchwork
                 )
source(here::here("scripts","utility.R"))

@@

list2env(stats, envir=environment())
```

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = "Figure 1"}
plot1()
```

@@

## Undetected proximodistal bias in TGI
Previous research demonstrated a phenomenon known as distal inhibition, wherein heat pain ratings tend to increase when a participant evaluates a more distal compared to a more proximal stimulus among two warm stimuli presented on the forearm [@quevedo_psychophysicshyperalgesia_2004; @quevedo_illusion_2007]. In our study, we sought to investigate whether this effect of distal inhibition would also influence TGI perception within a single dermatome (Fig. 3). Interestingly, we found no perceptual differences in TGI based on whether the cold stimulus was located proximally or distally (Exp. 1: xxx, xxx; Exp. 2: xxx, xxx). This result implies that the concept of distal inhibition may not apply when stimuli of different thermosensory qualities are presented. Alternatively, if it does apply, its effect size might be rather small, with limited impact on TGI perception.
<!-- Revision this but is proabably Run individual models on within and across dermatomes for H2 -->

```{r, results="hide", fig.width = 7.2, fig.height = 7.2, warning = F, message = F, echo = F, fig.cap = "Figure 2"}
plot2()
```

## Directional effects in inter-segmental sensory integration 
A main objective of this study was assessing spatial order effects along the rostrocaudal axis at the spinal level. We delivered an equal number of trials in which the cold stimulus was applied on a dermatome that mapped more rostrally or caudally compared to the warm or neutral stimuli. The results indicated a notably enhanced TGI effect when the cold stimulus induced more caudal activity within the spinal cord, as depicted in Figure 3. Our data revealed a substantial decrease in the subjective experience of cold (Exp. 1: estimate = xxx, p = xxx; Exp. 2: estimate: xxx, p = xxx), coupled with a significant increase in the subjective sensations of warmth (Exp. 1: estimate = xxx, p = xxx; Exp. 2: estimate: xxx, p = xxx) and burning sensations (Exp.1: estimate: xxx, p = xxx; Exp. 2 estimate: xxx, p = xxx) in response to TGI as compared to non-TGI stimuli. These findings were consistent across both experiments for thermosensory ratings, albeit with minor deviations that corresponded with the particular stimulation quality being assessed. The observed enhancement of warmth perception in Experiment 2 could be ascribed to the participants’ assessment of the warmer thermode as opposed to the colder one of Experiment 1. For burning ratings, significant results were seen in Experiment 2, but not in Experiment 1. One possible reason for this discrepancy could be the relatively low power in Experiment 1, as the power analysis was specifically focused on thermosensory ratings. Alternatively, assessing the warmer thermode might be a more accurate method for measuring TGI perception.

@@

## Participants
The study entailed two separate experiments, collectively involving 80 healthy volunteers. The sample consisted of xx females and xx males, mean age = XX years old (SD = x.xx, range = xx - xx) in Experiment 1, and xx females and xx males, mean age = XX years old (SD = x.xx, range = xx - xx), in Experiment 2. The research methodology complied with the principles set forth in the Declaration of Helsinki and received ethical approval from the Institutional Review Board (IRB) at the Danish Neuroscience Center, Aarhus University, Denmark. Prior to commencing the study, all participants were fully informed about the procedures and provided their voluntary consent.

@@

## Data analyses
We re-scaled data from cold, warm and burning VAS ratings from their original values to a range of 0 to 1. Following re-scaling, we applied zero-inflated mixed-effects beta regression models separately for each set of VAS ratings. In these models, we incorporated three fixed effects. These included the type of stimulation (non-TGI vs. TGI), the dermatome condition (within the same dermatome vs. across different dermatomes) and the spatial positioning of the cold or neutral thermode (proximal vs. distal within dermatomes; rostral vs. caudal  across dermatomes). These choices allowed us to assess the individual and interactive effects of these three factors on VAS ratings. Further, we added random intercepts to our models to account for between-subject variability and the effects of repeated measures. The variables introduced as random intercepts included the participant ID, the counterbalancing order and the trial number. The choices of the zero-inflated approach and the use of beta regressions were necessitated by the specific distribution of VAS ratings. The beta distribution is suitable for modelling VAS rating data, as they are proportional in nature. Additionally, the zero-inflation was needed due to the presence of an excess number of zero values in specific ratings and conditions. Specifically, we anticipated an overrepresentation of zero values for thermosensory ratings that were counterfactual to the objective stimulation quality (i.e., cold ratings of warm stimuli and warm ratings of cold stimuli) and burning ratings of non-TGI stimuli. The latter stimuli were designed to not elicit an illusion or trigger a weaker illusion as compared to the TGI stimuli. 
We carried out the statistical analyses using the 'glmmTMB' package in R (version x.xx), and statistical significance was set at p < .05. The experimental procedure, power analyses to determine sample size and statistical approach were preregistered for both Experiment 1 (add link) and Experiment 2 (add link). All data and code for the analysis are available in the github repository (xxx), ensuring the reproducibility of our findings.

>>>>

## Thermosensory and burning components of TGI perception are spinally mediated 
```{r, warning = F, message = F, echo = F,include=FALSE}
if (!require("pacman")) install.packages("pacman")
  pacman::p_load(boot, broom, broom.mixed, brms, car, caret, corrplot, DHARMa, gamlss,
                 ggeffects, gghalves, ggpol, ggpubr, glmmTMB, groupdata2, lme4, lmerTest, 
                 modelsummary, MuMIn, polycor, RColorBrewer, rcompanion, reshape2, Rmisc,
                 rsample, rstatix, simr, tibble, tidyverse, wesanderson, patchwork, cowplot
                 )
source(here::here("scripts","utility.R"))

@@

list2env(stats, envir=environment())

#save.image("~/tgi-spinal/Workspace/stats.RData")

```

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = "Figure 1"}
ggdraw() +
    draw_image(magick::image_read(here::here("figures","figure1.PNG")), scale = 1)
```

```{r, results="hide", fig.width = 7.2, fig.height = 5, warning = F, message = F, echo = F, fig.cap = "Figure 2"}
plot1()
```

@@

## Undetected proximodistal bias in TGI
Previous research demonstrated a phenomenon known as distal inhibition, wherein heat pain ratings tend to increase when a participant evaluates a more distal compared to a more proximal stimulus among two warm stimuli presented on the forearm [@quevedo_psychophysicshyperalgesia_2004; @quevedo_illusion_2007]. In our study, we sought to investigate whether this effect of distal inhibition would also influence TGI perception within a single dermatome (Fig. 3).
Firstly the cold intensity was less when the cold probe was located more distally for both control and TGI trials (TGI: $\beta$ = `r stats_within_cold_exp1_TGI$beta[2]`, `r make_pvalue(stats_within_cold_exp1_TGI$p[2])`; control: $\beta$ = `r stats_within_cold_exp1_control$beta[2]`, `r make_pvalue(stats_within_cold_exp1_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_within_cold_exp1_control$beta[5]`, `r make_pvalue(stats_within_cold_exp1_control$p[5])`. No difference in intensity was found for the warm ratings (TGI: $\beta$ = `r stats_within_warm_exp1_TGI$beta[2]`, `r make_pvalue(stats_within_warm_exp1_TGI$p[2])`; control: $\beta$ = `r stats_within_warm_exp1_control$beta[2]`, `r make_pvalue(stats_within_warm_exp1_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_within_warm_exp1_control$beta[5]`, `r make_pvalue(stats_within_warm_exp1_control$p[5])`. significant differences was also found for the burning ratings indicating more distally located cold probe decreased ratings. (TGI: $\beta$ = `r stats_within_burn_exp1_TGI$beta[2]`, `r make_pvalue(stats_within_burn_exp1_TGI$p[2])`; control: $\beta$ = `r stats_within_burn_exp1_control$beta[2]`, `r make_pvalue(stats_within_burn_exp1_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_within_burn_exp1_control$beta[5]`, `r make_pvalue(stats_within_burn_exp1_control$p[5])`.

The second experiment showed similar results 
Firstly the cold intensity was less when the cold probe was located more distally for both control and TGI trials (TGI: $\beta$ = `r stats_within_cold_exp2_TGI$beta[2]`, `r make_pvalue(stats_within_cold_exp2_TGI$p[2])`; control: $\beta$ = `r stats_within_cold_exp2_control$beta[2]`, `r make_pvalue(stats_within_cold_exp2_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_within_cold_exp2_control$beta[5]`, `r make_pvalue(stats_within_cold_exp2_control$p[5])`. No difference in intensity was found for the warm ratings (TGI: $\beta$ = `r stats_within_warm_exp2_TGI$beta[2]`, `r make_pvalue(stats_within_warm_exp2_TGI$p[2])`; control: $\beta$ = `r stats_within_warm_exp2_control$beta[2]`, `r make_pvalue(stats_within_warm_exp2_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_within_warm_exp2_control$beta[5]`, `r make_pvalue(stats_within_warm_exp2_control$p[5])`. significant differences was also found for the burning ratings indicating more distally located cold probe decreased ratings. (TGI: $\beta$ = `r stats_within_burn_exp2_TGI$beta[2]`, `r make_pvalue(stats_within_burn_exp2_TGI$p[2])`; control: $\beta$ = `r stats_within_burn_exp2_control$beta[2]`, `r make_pvalue(stats_within_burn_exp2_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_within_burn_exp2_control$beta[5]`, `r make_pvalue(stats_within_burn_exp2_control$p[5])`.

```{r, results="hide", fig.width = 7.2, fig.height = 7.2, warning = F, message = F, echo = F, fig.cap = "Figure 3"}
plot2()
```

## Directional effects in inter-segmental sensory integration 
A main objective of this study was assessing spatial order effects along the rostrocaudal axis at the spinal level. We delivered an equal number of trials in which the cold stimulus was applied on a dermatome that mapped more rostrally or caudally compared to the warm or neutral stimuli.

Firstly the cold intensity was less when the cold probe was located more distally for both control and TGI trials (TGI: $\beta$ = `r stats_across_cold_exp1_TGI$beta[2]`, `r make_pvalue(stats_across_cold_exp1_TGI$p[2])`; control: $\beta$ = `r stats_across_cold_exp1_control$beta[2]`, `r make_pvalue(stats_across_cold_exp1_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_across_cold_exp1_control$beta[5]`, `r make_pvalue(stats_across_cold_exp1_control$p[5])`. No difference in intensity was found for the warm ratings (TGI: $\beta$ = `r stats_across_warm_exp1_TGI$beta[2]`, `r make_pvalue(stats_across_warm_exp1_TGI$p[2])`; control: $\beta$ = `r stats_across_warm_exp1_control$beta[2]`, `r make_pvalue(stats_across_warm_exp1_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_across_warm_exp1_control$beta[5]`, `r make_pvalue(stats_across_warm_exp1_control$p[5])`. significant differences was also found for the burning ratings indicating more distally located cold probe decreased ratings. (TGI: $\beta$ = `r stats_across_burn_exp1_TGI$beta[2]`, `r make_pvalue(stats_across_burn_exp1_TGI$p[2])`; control: $\beta$ = `r stats_across_burn_exp1_control$beta[2]`, `r make_pvalue(stats_across_burn_exp1_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_across_burn_exp1_control$beta[5]`, `r make_pvalue(stats_across_burn_exp1_control$p[5])`.

The second experiment showed similar results 
Firstly the cold intensity was less when the cold probe was located more distally for both control and TGI trials (TGI: $\beta$ = `r stats_across_cold_exp2_TGI$beta[2]`, `r make_pvalue(stats_across_cold_exp2_TGI$p[2])`; control: $\beta$ = `r stats_across_cold_exp2_control$beta[2]`, `r make_pvalue(stats_across_cold_exp2_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_across_cold_exp2_control$beta[5]`, `r make_pvalue(stats_across_cold_exp2_control$p[5])`. No difference in intensity was found for the warm ratings (TGI: $\beta$ = `r stats_across_warm_exp2_TGI$beta[2]`, `r make_pvalue(stats_across_warm_exp2_TGI$p[2])`; control: $\beta$ = `r stats_across_warm_exp2_control$beta[2]`, `r make_pvalue(stats_across_warm_exp2_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_across_warm_exp2_control$beta[5]`, `r make_pvalue(stats_across_warm_exp2_control$p[5])`. significant differences was also found for the burning ratings indicating more distally located cold probe decreased ratings. (TGI: $\beta$ = `r stats_across_burn_exp2_TGI$beta[2]`, `r make_pvalue(stats_across_burn_exp2_TGI$p[2])`; control: $\beta$ = `r stats_across_burn_exp2_control$beta[2]`, `r make_pvalue(stats_across_burn_exp2_control$p[2])`) with no significant interaction between them  $\beta$ = `r stats_across_burn_exp2_control$beta[5]`, `r make_pvalue(stats_across_burn_exp2_control$p[5])`.

The results indicated a notably enhanced TGI effect when the cold stimulus induced more caudal activity within the spinal cord, as depicted in Figure 3. These findings were consistent across both experiments for thermosensory ratings, albeit with minor deviations that corresponded with the particular stimulation quality being assessed. The observed enhancement of warmth perception in Experiment 2 could be ascribed to the participants’ assessment of the warmer thermode as opposed to the colder one of Experiment 1. For burning ratings, significant results were seen in Experiment 2, but not in Experiment 1. One possible reason for this discrepancy could be the relatively low power in Experiment 1, as the power analysis was specifically focused on thermosensory ratings. Alternatively, assessing the warmer thermode might be a more accurate method for measuring TGI perception.

@@

## Participants
The study entailed two separate experiments, collectively involving 80 healthy volunteers. The sample consisted of `r descript_exp1_gender %>% filter(Gender == "F") %>% .$number` females and `r descript_exp1_gender %>% filter(Gender == "M") %>% .$number` males, mean age = `r descript_exp1_age$mean_age` years old (SD = `r descript_exp1_age$sd_age`, range = `r descript_exp1_age$min_age` - `r descript_exp1_age$max_age`) in Experiment 1,

and `r descript_exp2_gender %>% filter(Gender == "F") %>% .$number` females and `r descript_exp2_gender %>% filter(Gender == "M") %>% .$number` males and `r descript_exp2_gender %>% filter(Gender != "M" & Gender != "F") %>% .$number` non-binary (Female at birth), mean age = `r descript_exp2_age$mean_age` years old (SD = `r descript_exp2_age$sd_age`, range = `r descript_exp2_age$min_age` - `r descript_exp2_age$max_age`), in Experiment 2. The research methodology complied with the principles set forth in the Declaration of Helsinki and received ethical approval from the Institutional Review Board (IRB) at the Danish Neuroscience Center, Aarhus University, Denmark. Prior to commencing the study, all participants were fully informed about the procedures and provided their voluntary consent.

@@

## Data analyses
We re-scaled data from cold, warm and burning VAS ratings from their original values to a range of 0 to 1. Following re-scaling, we applied zero-inflated mixed-effects beta regression models separately for each set of VAS ratings. In these models, we incorporated three fixed effects. These included the type of stimulation (non-TGI vs. TGI), the dermatome condition (within the same dermatome vs. across different dermatomes) and the spatial positioning of the cold or neutral thermode (proximal vs. distal within dermatomes; rostral vs. caudal  across dermatomes). These choices allowed us to assess the individual and interactive effects of these three factors on VAS ratings. Further, we added random intercepts to our models to account for between-subject variability and the effects of repeated measures. The variables introduced as random intercepts included the participant ID, the counterbalancing order and the trial number. The choices of the zero-inflated approach and the use of beta regressions were necessitated by the specific distribution of VAS ratings. The beta distribution is suitable for modelling VAS rating data, as they are proportional in nature. Additionally, the zero-inflation was needed due to the presence of an excess number of zero values in specific ratings and conditions. Specifically, we anticipated an overrepresentation of zero values for thermosensory ratings that were counterfactual to the objective stimulation quality (i.e., cold ratings of warm stimuli and warm ratings of cold stimuli) and burning ratings of non-TGI stimuli. The latter stimuli were designed to not elicit an illusion or trigger a weaker illusion as compared to the TGI stimuli. 
We carried out the statistical analyses using the 'glmmTMB' package in R (version `r sessionInfo()$otherPkgs$glmmTMB$Version`), and statistical significance was set at p < .05. The experimental procedure, power analyses to determine sample size and statistical approach were preregistered for both [Experiment 1](https://osf.io/4xcn5/) and [Experiment 2](https://osf.io/dhg8u/). All data and code for the analysis are available in the [github repository](https://github.com/Body-Pain-Perception-Lab/tgi-spinal/tree/Markdown-manuscript), ensuring the reproducibility of our findings.

### 2023-07-12 tgi-spinal_analysis-exp1.Rmd
updates

source(here::here("scripts","utility.R"))
```

@@

summary_stats(model.cold, 3,2)
Anova(model.cold)
# get Rsq with the mumin package
```
# Marginal effect plots of full model - cold
```{r}
cold_eff <- ggeffects::ggpredict(model.cold, terms = c('cold_cond', 'manipulation', 'condition'),
                           ci.lvl = .95)
# changing labels for plotting
cold_eff$group <- factor(cold_eff$group, labels = c('Non-TGI', 'TGI'))
cold_eff <- cold_eff %>% 
  mutate(label = 
           case_when(facet == 'across' & x == 'prox_caud' ~ 'caudal',
                     facet == 'across' & x == 'dist_rostr' ~ 'rostral',
                     facet == 'within' & x == 'prox_caud' ~ 'proximal',
                     facet == 'within' & x == 'dist_rostr' ~ 'distal',) 
         )

>>>>

source(here::here("scripts","utility.R"))
source(here::here("scripts","models.R"))

@@

summary_stat(model.cold, 3,2)
Anova(model.cold)
# get Rsq with the mumin package
```
# Marginal effect plots of full model - cold
```{r}
cold_eff <- ggeffects::ggpredict(model.cold, terms = c('cold_cond', 'manipulation', 'condition'),
                           ci.lvl = .95)
# changing labels for plotting
cold_eff$group <- factor(cold_eff$group, labels = c('Non-TGI', 'TGI'))
cold_eff <- cold_eff %>% 
  mutate(label = 
           case_when(facet == 'across' & x == 'prox_caud' ~ 'caudal',
                     facet == 'across' & x == 'dist_rostr' ~ 'rostral',
                     facet == 'within' & x == 'prox_caud' ~ 'proximal',
                     facet == 'within' & x == 'dist_rostr' ~ 'distal',) 
         )
